---
title: "Effet des annonces Macron"
author: "FD"
output: 
  html_document: 
      code_folding: hide
      toc: TRUE
      toc_float: TRUE
      self_contained: no
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

```{r, include = FALSE}
# File names
dataFileCom <- paste0("data/donnees-de-vaccination-par-commune_2021-08-06.csv") 
dataFileEPCI <- paste0("data/donnees-de-vaccination-par-epci_2021-08-06.csv")
```

```{r, results = 'hide', warning=FALSE}
library(plotly)
library(RColorBrewer)
```

# Initializations

## Load data

### Niveau de vie par EPCI (2018)

Source <https://www.insee.fr/fr/statistiques/5009236?sommaire=5009255&q=revenu+epci#dictionnaire>

<!-- 
Code géographique ;
Libellé géographique ;
Nombre de ménages fiscaux ;
Nombre de personnes dans les ménages fiscaux ;
Médiane du niveau de vie (€) ;
Part des ménages fiscaux imposés (%) ;
Taux de pauvreté-Ensemble (%) ;
Taux de pauvreté des ménages dont le référent fiscal a moins de 30 ans (%) ;
Taux de pauvreté des ménages dont le référent fiscal a de 30 à 39 ans (%) ;
Taux de pauvreté des ménages dont le référent fiscal a de 40 à 49 ans (%) ;
Taux de pauvreté des ménages dont le référent fiscal a de 50 à 59 ans (%) ;
Taux de pauvreté des ménages dont le référent fiscal a de 60 à 74 ans (%) ;
Taux de pauvreté des ménages dont le référent fiscal a 75 ans ou plus (%) ;
Taux de pauvreté des ménages propriétaires de leur logement (%) ;
Taux de pauvreté des ménages locataires de leur logement (%) ;
Part des revenus d'activité (%) ;
dont part des salaires, traitements (%) ;
dont part des indemnités de chômage (%) ;
dont part des revenus des activités non salariées (%) ;
Part des pensions, retraites et rentes (%) ;
Part des revenus du patrimoine et des autres revenus (%) ;
Part de l'ensemble des prestations sociales (%) ;
dont part des prestations familiales (%) ;
dont part des minima sociaux (%) ;
dont part des prestations logement (%) ;
Part des impôts (%) ;
Rapport interdécile 9e décile/1er decile ;
1er décile du niveau de vie (€) ;
9e décile du niveau de vie (€).
-->

```{r loadDataRevenu, results = 'hide'}
dat.revenuEPCI <- read.csv("data/insee/cc_filosofi_2018_EPCI-geo2021.CSV", sep = ";")
head(dat.revenuEPCI)

dat.revenuCom <- read.csv("data/cc_filosofi_2018_COM-geo2021.csv", sep = ";")
head(dat.revenuCom)
```


### Vaccination par EPCI et par commune

Source <https://datavaccin-covid.ameli.fr/explore/dataset/donnees-de-vaccination-par-epci/download/?format=csv&timezone=Europe/Berlin&lang=fr&use_labels_for_header=true&csv_separator=%3B>

```{r, results = 'hide'} 
dat.vaccinationEPCI <- read.csv(dataFileEPCI, sep = ";", stringsAsFactors = FALSE, dec = ",")
head(dat.vaccinationEPCI)

dat.vaccinationCom <- read.csv(dataFileCom, sep = ";", stringsAsFactors = FALSE, dec = ",")
head(dat.vaccinationCom)
```

### Composition communales 

Pour les infos sur les départements

Source <https://www.insee.fr/fr/information/2510634>
```{r}
composition <- read.csv("data/EPCI_composition-communale.csv", encoding = "UFT-8")
```

Source region codes <https://www.data.gouv.fr/en/datasets/regions-de-france/>

```{r}
# Departements
departements <- read.csv("data/departement2020.csv")

dic.depname <- departements$libelle
names(dic.depname) <- departements$dep

# Load region codes 
reg <- read.csv("data/regions-france.csv")
```

## Clean data

```{r}
# Final week in the data
maxWeek1 <- max(unique(dat.vaccinationEPCI$semaine_injection))
maxWeek2 <- max(unique(dat.vaccinationCom$semaine_injection))
stopifnot(maxWeek1 == maxWeek2)
maxWeek <- maxWeek1
```

```{r, results='hide'}
# Weeks in the datasets
wks1 <- sort(unique(dat.vaccinationEPCI$semaine_injection))
wks2 <- sort(unique(dat.vaccinationCom$semaine_injection))
stopifnot(all(wks1 == wks2))
# Check that consecutive weeks
stopifnot(all(is.element(diff(as.numeric(substr(wks1, 6, 7))), c(1, -52))))
wks <- wks1

wks[1] # Show first week; week 52 of 2020 started on 2020-12-21
# Dates of Mondays of the corresponding weeks
beginWeek <- seq(as.Date("2020-12-21"), as.Date("2021-12-31"), by = 7)
names(beginWeek) <- wks
# Dates of Sundays of the corresponding weeks
endWeek <- as.Date(beginWeek) + 6
names(endWeek) <- wks

# Add the day infos to the datasets
dat.vaccinationEPCI$beginWeek <- beginWeek[dat.vaccinationEPCI$semaine_injection]
dat.vaccinationEPCI$endWeek <- endWeek[dat.vaccinationEPCI$semaine_injection]
dat.vaccinationEPCI$week.DD <- paste(dat.vaccinationEPCI$beginWeek, dat.vaccinationEPCI$endWeek, sep = "_")

dat.vaccinationCom$beginWeek <- beginWeek[dat.vaccinationCom$semaine_injection]
dat.vaccinationCom$endWeek <- endWeek[dat.vaccinationCom$semaine_injection]
dat.vaccinationCom$week.DD <- paste(dat.vaccinationCom$beginWeek, dat.vaccinationCom$endWeek, sep = "_")
```

Merge the vaccination and revenu datasets (all times)

```{r, results = 'hide'}
# First, compare contents
## EPCI
# Number of different EPCI in the two datasets
c(vaccin = length(unique(dat.vaccinationEPCI$epci)), revenu = length(unique(dat.revenuEPCI$CODGEO)))

# Compare EPCIs
# Some vaccin are not included in revenu
any(!is.element(unique(dat.vaccinationEPCI$epci), unique(dat.revenuEPCI$CODGEO)))

# All revenu are included in vaccin
any(!is.element(unique(dat.revenuEPCI$CODGEO), unique(dat.vaccinationEPCI$epci)))

## Communes
# Check if vaccination communes included in revenu dataset
any(!is.element(unique(dat.vaccinationCom$commune_residence), unique(dat.revenuCom$CODGEO)))
# Show them -- codes when location is unknown
unique(dat.vaccinationCom$commune_residence)[!is.element(unique(dat.vaccinationCom$commune_residence), unique(dat.revenuCom$CODGEO))]


# Merge the datasets 
## EPCI
dat.EPCI <- merge(dat.vaccinationEPCI, dat.revenuEPCI, by.x = "epci", by.y = "CODGEO", all.x = TRUE)
stopifnot(nrow(dat.vaccinationEPCI) == nrow(dat.EPCI)) # Check all included

## Communes
dat.Com <- merge(dat.vaccinationCom, dat.revenuCom, by.x = "commune_residence", by.y = "CODGEO", all.x = TRUE)
stopifnot(nrow(dat.vaccinationCom) == nrow(dat.Com)) # Check all included
```


Departement information

```{r, results='hide'}
# Information about departement in which the different EPCI are
# Some are across multiple departements: keep the information by collating them with "_"
agg_nbdep <- aggregate(composition$DEP, by = list(composition$EPCI), FUN = function(i) paste(sort(unique(i)), collapse = "_"))
table(agg_nbdep$x)

# Dictionnary of departement(s) associated to EPCI
dic.dep <- agg_nbdep$x
names(dic.dep) <- agg_nbdep$Group.1

# Add the dep information to our EPCI data
dat.EPCI$dep <- dic.dep[as.character(dat.EPCI$epci)]

# Add the libelle information
dat.EPCI$dep_libelle <- dic.depname[as.character(dat.EPCI$dep)]

# Get departement information in Communes dataset
dat.Com$dep <- substr(dat.Com$commune_residence, 1, 2)
```

Add region information 

```{r, results = 'hide'}
# As dictionnary
reg.dic <- c(reg$nom_region)
names(reg.dic) <- reg$code_region # !! Name has to be in quotes

# Initialize region data
unique(dat.EPCI$reg_code)

# Some EPCIs are on two regions
dat.EPCI$twoRegions <- (nchar(dat.EPCI$reg_code) > 2)

# Extract regional code, including for EPCI on two regions
dat.EPCI$reg_code1 <- substr(dat.EPCI$reg_code, 1, 2)
dat.EPCI$reg_code2 <- substr(dat.EPCI$reg_code, 4, 5)
dat.EPCI[which(dat.EPCI$reg_code2 == ""), "reg_code2"] <- NA

dat.EPCI$libelle_region <- reg.dic[dat.EPCI$reg_code1]
dat.EPCI$libelle_region2 <- reg.dic[dat.EPCI$reg_code2]

# Add short names
# Add shorter name
dat.EPCI$reg_shortname <- NA
dat.EPCI[which(dat.EPCI$libelle_region == "Île-de-France"), "reg_shortname"] <- "IDF"
dat.EPCI[which(dat.EPCI$libelle_region == "Provence-Alpes-Côte d'Azur"), "reg_shortname"] <- "PACA"
dat.EPCI[which(dat.EPCI$libelle_region == "Bretagne"), "reg_shortname"] <- "BRE"
dat.EPCI[which(dat.EPCI$libelle_region == "Auvergne-Rhône-Alpes"), "reg_shortname"] <- "ARA"
dat.EPCI[which(dat.EPCI$libelle_region == "Hauts-de-France"), "reg_shortname"] <- "HDF"
dat.EPCI[which(dat.EPCI$libelle_region == "Grand Est"), "reg_shortname"] <- "GE"
dat.EPCI[which(dat.EPCI$libelle_region == "Occitanie"), "reg_shortname"] <- "OCC"
dat.EPCI[which(dat.EPCI$libelle_region == "Normandie"), "reg_shortname"] <- "NOR"
dat.EPCI[which(dat.EPCI$libelle_region == "Bourgogne-Franche-Comté"), "reg_shortname"] <- "BFC"
dat.EPCI[which(dat.EPCI$libelle_region == "Pays de la Loire"), "reg_shortname"] <- "PDL"
dat.EPCI[which(dat.EPCI$libelle_region == "Nouvelle-Aquitaine"), "reg_shortname"] <- "NAQ"
dat.EPCI[which(dat.EPCI$libelle_region == "Centre-Val de Loire"), "reg_shortname"] <- "CVL"
dat.EPCI[which(dat.EPCI$libelle_region == "Corse"), "reg_shortname"] <- "COR"

table(dat.EPCI$reg_shortname, useNA = "ifany")
table(dat.EPCI$libelle_region, useNA = "ifany")

# Missing DOMs

## Coms
unique(dat.Com$dep)
# Initialize region data
dat.Com$reg_shortname <- NA
dat.Com$region <- NA
# Get region from departement information
dat.Com[base::is.element(dat.Com$dep, c("13", "83", "84")), c("reg_shortname", "region")] <- c("PACA", "Provence-Alpes-Côte d'Azur")
dat.Com[base::is.element(dat.Com$dep, c("75", "91", "92", "93", "94", "95")), c("reg_shortname", "region")] <- c("IDF", "Île-de-France")
dat.Com[base::is.element(dat.Com$dep, c("69")), c("reg_shortname", "region")] <- c("ARA", "Auvergne-Rhône-Alpes")
```

Make sure that values are numeric 

```{r, results = 'hide'}
for(col in c("taux_1_inj", "taux_termine", 
             "taux_cumu_1_inj", "taux_cumu_termine")){
  dat.Com[, col] <- as.numeric(dat.Com[, col])
  dat.EPCI[, col] <- as.numeric(dat.EPCI[, col])
}
```

Dictionary of age classes

```{r, results = 'hide'}
# Check that age classes match
all(sort(unique(dat.Com$classe_age)) == sort(unique(dat.EPCI$classe_age)))
dic.ages <- as.character(unique(dat.Com$classe_age))
names(dic.ages) <- unique(dat.Com$libelle_classe_age)
dic.ages
```


## Plot settings

Region colors and pch

```{r}
# Define colors, joining palettes
manycols <- c(brewer.pal(name = "Set2", 8), brewer.pal(name = "Set1", 8))
# Region colors
colRegion <- manycols[1:length(unique(dat.EPCI$reg_shortname))]
names(colRegion) <- unique(dat.EPCI$reg_shortname)

# Region pch
pchRegion <- 14 + 1:length(unique(dat.EPCI$reg_shortname))
names(pchRegion) <- names(colRegion)
```

Age colors 

```{r}
ages <- unique(dat.EPCI$classe_age)
colAge <- brewer.pal(name = "Dark2", n = length(ages))
names(colAge) <- sort(ages)
```

# Some checks

Compare MED18 and PIMP18

```{r}
tmp <- dat.EPCI[which(dat.EPCI$classe_age == "TOUT_AGE" & dat.EPCI$semaine_injection == maxWeek), ]
plot(tmp$PIMP18, tmp$MED18, xlab = "PIMP18, Part de ménages fiscaux imposés", ylab = "MED18, revenu médian")

tmp <- dat.Com[which(dat.Com$classe_age == "TOUT_AGE" & dat.Com$semaine_injection == maxWeek), ]
plot(tmp$PIMP18, tmp$MED18, xlab = "PIMP18, Part de ménages fiscaux imposés", ylab = "MED18, revenu médian")
```


# Model and plot

## Final week only

### All regions together, PIMP18

```{r}
mdlFinal.1D.PIMP.0 <- glm(cbind(effectif_1_inj, population_carto - effectif_1_inj) ~ PIMP18 + as.factor(classe_age) + PIMP18 * as.factor(classe_age) + PIMP18 * as.factor(reg_shortname) + as.factor(reg_shortname), family = binomial(link = "logit"), data = dat.EPCI[which(dat.EPCI$classe_age != "TOUT_AGE" & dat.EPCI$semaine_injection == maxWeek), ])


mdlFinal.1D.PIMP.noregion <- glm(cbind(effectif_1_inj, population_carto - effectif_1_inj) ~ PIMP18 + as.factor(classe_age) + PIMP18 * as.factor(classe_age), family = binomial(link = "logit"), data = dat.EPCI[which(dat.EPCI$classe_age != "TOUT_AGE" & dat.EPCI$semaine_injection == maxWeek), ])

mdlComFinal.1D.PIMP.noregion <- glm(cbind(effectif_1_inj, population_carto - effectif_1_inj) ~ PIMP18 + as.factor(classe_age) + PIMP18 * as.factor(classe_age), family = binomial(link = "logit"), data = dat.Com[which(dat.Com$classe_age != "TOUT_AGE" & dat.Com$semaine_injection == maxWeek), ])

mdlFinal.termine.PIMP.noregion <- glm(cbind(effectif_termine, population_carto - effectif_termine) ~ PIMP18 + as.factor(classe_age) + PIMP18 * as.factor(classe_age), family = binomial(link = "logit"), data = dat.EPCI[which(dat.EPCI$classe_age != "TOUT_AGE" & dat.EPCI$semaine_injection == maxWeek), ])

summary(mdlFinal.1D.PIMP.0)
car::Anova(mdlFinal.1D.PIMP.0)

summary(mdlFinal.1D.PIMP.noregion)
car::Anova(mdlFinal.1D.PIMP.noregion)

summary(mdlComFinal.1D.PIMP.noregion)
car::Anova(mdlComFinal.1D.PIMP.noregion)

summary(mdlFinal.termine.PIMP.noregion)
car::Anova(mdlFinal.termine.PIMP.noregion)

```

### France entiere

#### Une dose au moins, EPCI

```{r plotFranceEntiereEPCI, fig.height=6.5}
par(xpd = TRUE, las = 1)
par(mar = c(4, 4, 4, 4))
par(mgp = c(2, 0.5, 0), tck = -0.01)

tmpAll <- dat.EPCI[which(dat.EPCI$classe_age != "TOUT_AGE" & dat.EPCI$semaine_injection == maxWeek), ]
themaxpop <- max(tmpAll$population_carto)
tmpAll$relsize <- tmpAll$population_carto / themaxpop
tmpAll$relsize.transfo <- 5*(tmpAll$relsize*3)^(1/3)

plot(tmpAll$PIMP18, 100*tmpAll$taux_1_inj, 
     col = adjustcolor(colAge[tmpAll$classe_age], 0.5), 
     pch = 16, cex = tmpAll$relsize.transfo,
     ylim = c(0, 10), xlim = c(15, 85), 
     frame.plot = FALSE, yaxs = "i", #xaxs = "i", 
     xlab = "Part des ménages fiscaux imposés dans l'EPCI en 2018 (%)", ylab = "Taux de vaccination 1 dose au moins (%)",
     type = "n"
     )
par(xpd = FALSE)
for(i in seq(0, 10, by = 1)){
  abline(h = i, col = gray(0.8), lty = 1)
}

legend("topleft", col = colAge[-length(colAge)], legend = names(colAge[-length(colAge)]), pch = 16, bty = "n")
axis(4)


points(tmpAll$PIMP18, 100*tmpAll$taux_1_inj, 
     col = adjustcolor(colAge[tmpAll$classe_age], 0.5), 
     pch = 16, cex = tmpAll$relsize.transfo,
)
newdata.noregion <- expand.grid(PIMP18 = seq(min(dat.EPCI$PIMP18, na.rm = TRUE), max(dat.EPCI$PIMP18, na.rm = TRUE), length.out = 100), classe_age = names(colAge[-length(colAge)]))

ndt.noregion <- newdata.noregion
ndt.noregion$prd1D <- predict(mdlFinal.1D.PIMP.noregion, newdata = newdata.noregion, type = "response")

lwd.pred <- 2
for(age in ages[ages!="TOUT_AGE"]){
  # Get predicted data for this age class (no region)
  subd <- ndt.noregion[ndt.noregion$classe_age == age,]
  lines(subd$PIMP18, 100*subd$prd1D, col = colAge[age], lwd = lwd.pred)
}

convertDate <- function(x) format.Date(as.Date(x), "%d/%m")
mtext(side = 3, paste0("France entière\nPendant la semaine du ", convertDate(unique(tmpAll$beginWeek)), " au ", convertDate(unique(tmpAll$endWeek))), line = 1.)

```


#### Une dose au moins, Communes

```{r plotFranceEntiereCommunes, fig.height=6.5}
par(xpd = TRUE, las = 1)
par(mar = c(4, 4, 4, 4))
par(mgp = c(2, 0.5, 0), tck = -0.01)

tmpAll <- dat.Com[which(dat.Com$classe_age != "TOUT_AGE" & dat.Com$semaine_injection == maxWeek), ]
themaxpop <- max(tmpAll$population_carto)
tmpAll$relsize <- tmpAll$population_carto / themaxpop
tmpAll$relsize.transfo <- 2*(tmpAll$relsize*3)^(1/3)

plot(tmpAll$PIMP18, 100*tmpAll$taux_1_inj, 
     col = adjustcolor(colAge[tmpAll$classe_age], 0.5), 
     pch = 16, cex = tmpAll$relsize.transfo,
     ylim = c(0, 10), xlim = c(15, 85), 
     frame.plot = FALSE, yaxs = "i", #xaxs = "i", 
     xlab = "Part des ménages fiscaux imposés dans la commune en 2018 (%)", ylab = "Taux de vaccination 1 dose au moins (%)",
     type = "n"
     )
par(xpd = FALSE)
for(i in seq(0, 10, by = 1)){
  abline(h = i, col = gray(0.8), lty = 1)
}

legend("topleft", col = colAge[-length(colAge)], legend = names(colAge[-length(colAge)]), pch = 16, bty = "n")
axis(4)


points(tmpAll$PIMP18, 100*tmpAll$taux_1_inj, 
     col = adjustcolor(colAge[tmpAll$classe_age], 0.5), 
     pch = 16, cex = tmpAll$relsize.transfo,
)
newdata.noregion <- expand.grid(PIMP18 = seq(min(dat.Com$PIMP18, na.rm = TRUE), max(dat.Com$PIMP18, na.rm = TRUE), length.out = 100), classe_age = names(colAge[-length(colAge)]))

ndt.noregion <- newdata.noregion
ndt.noregion$prd1D <- predict(mdlComFinal.1D.PIMP.noregion, newdata = newdata.noregion, type = "response")

lwd.pred <- 2
for(age in ages[ages!="TOUT_AGE"]){
  # Get predicted data for this age class (no region)
  subd <- ndt.noregion[ndt.noregion$classe_age == age,]
  lines(subd$PIMP18, 100*subd$prd1D, col = colAge[age], lwd = lwd.pred)
}

convertDate <- function(x) format.Date(as.Date(x), "%d/%m")
mtext(side = 3, paste0("Communes autour de Paris, Marseille, Lyon\nPendant la semaine du ", convertDate(unique(tmpAll$beginWeek)), " au ", convertDate(unique(tmpAll$endWeek))), line = 1.)


```



