---
title: "Campagne vaccinale"
author: "FD"
output: 
  html_document: 
      code_folding: hide
      toc: TRUE
      toc_float: TRUE
      self_contained: no
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls()) # I don't care, I am doing it anyway
```

```{r, include = FALSE}
dlData <- FALSE
# Whether to download the data again
```


# Load data

## England 

Source England data
<https://coronavirus.data.gov.uk/details/download>

```{r}
URL.EN <- "https://api.coronavirus.data.gov.uk/v2/data?areaType=nation&areaCode=E92000001&metric=vaccinationsAgeDemographics&format=csv"
destEN <- "data/vaccAgeEN.csv"
  
if(dlData){
  download.file(URL.EN, destEN)
}

datEN <- read.csv(destEN)
```

## France

Source France data
<https://www.data.gouv.fr/en/datasets/donnees-relatives-aux-personnes-vaccinees-contre-la-covid-19-1/>

```{r}
URL.FR <- "https://www.data.gouv.fr/fr/datasets/r/54dd5f8d-1e2e-4ccb-8fb8-eac68245befd"
destFR <- "data/vaccAgeFR.csv"
if(dlData){
  download.file(URL.FR, destFR)
}

datFR <- read.csv(destFR, sep = ";")
```

# Clean and homogenize data

## Ages 

```{r}
maxAge <- 100

# EN
unique(datEN$age)

# Extract min and max age
datEN$ageMin <- substr(datEN$age, 1, 2)
datEN$ageMax <- substr(datEN$age, 4, 5)
datEN[datEN$age == "90+", "ageMax"] <- as.character(maxAge)

# Turn into numeric
datEN$ageMin <- as.numeric(datEN$ageMin)
datEN$ageMax <- as.numeric(datEN$ageMax)


# FR
# Remove all ages
datFR <- datFR[datFR$clage_vacsi > 0, ]
sort(unique(datFR$clage_vacsi))

datFR$ageMax <- datFR$clage_vacsi
datFR[datFR$ageMax == 80, "ageMax"] <- maxAge

datFR$ageMin <- NA
datFR[datFR$clage_vacsi == 4, "ageMin"] <- 0
datFR[datFR$clage_vacsi == 9, "ageMin"] <- 5
datFR[datFR$clage_vacsi == 11, "ageMin"] <- 10
datFR[datFR$clage_vacsi == 17, "ageMin"] <- 12
datFR[datFR$clage_vacsi == 24, "ageMin"] <- 18
datFR[datFR$clage_vacsi == 29, "ageMin"] <- 25
datFR[datFR$clage_vacsi == 39, "ageMin"] <- 30
datFR[datFR$clage_vacsi == 49, "ageMin"] <- 40
datFR[datFR$clage_vacsi == 59, "ageMin"] <- 50
datFR[datFR$clage_vacsi == 64, "ageMin"] <- 60
datFR[datFR$clage_vacsi == 69, "ageMin"] <- 65
datFR[datFR$clage_vacsi == 74, "ageMin"] <- 70
datFR[datFR$clage_vacsi == 79, "ageMin"] <- 75
datFR[datFR$clage_vacsi == 80, "ageMin"] <- 80

table(datFR$ageMax, useNA = "ifany")
table(datFR$ageMin, useNA = "ifany")

```

## Doses

Note: In France, people who have been previously infected skipped the second dose, and they were immediately "fully vaccinated" after their first dose. 

```{r}
datEN$d1 <- datEN$newPeopleVaccinatedFirstDoseByVaccinationDate
datEN$d2 <- datEN$newPeopleVaccinatedSecondDoseByVaccinationDate
datEN$d3 <- datEN$newPeopleVaccinatedThirdInjectionByVaccinationDate

datFR$d1 <- datFR$n_dose1
datFR$d2 <- datFR$n_complet
datFR$d3 <- datFR$n_rappel
```

## Dates

```{r}
# as date
datEN$date <- as.Date(datEN$date)
datFR$date <- as.Date(datFR$jour)


# Date range
range(c(datEN$date, datFR$date))
firstDate <- as.Date("2020-12-07") # Monday
mondays <- seq(firstDate, max(c(datEN$date, datFR$date))+7, by = "week")
mondays

# Number weeks from first week in the data
datEN$wk <- sapply(datEN$date, FUN = function(i) sum(i >= mondays))
datFR$wk <- sapply(datFR$date, FUN = function(i) sum(i >= mondays))

tmp <- sapply(mondays, FUN = function(i) sum(i >= mondays))
dates <- data.frame(wk = tmp, mondays = mondays, timeSinceFirstDay = as.numeric(mondays - firstDate))


# Add year and week
datEN$year <- format(datEN$date, "%Y")
datEN$week <- format(datEN$date, "%W")
datEN$month <- format(datEN$date, "%m")

datFR$year <- format(datFR$date, "%Y")
datFR$week <- format(datFR$date, "%W")
datFR$month <- format(datFR$date, "%m")

# Round days
months <- seq(as.Date("2020-12-01"), max(c(datFR$date, datEN$date)) + 30, by = "month")
tmp <- sapply(months, FUN = function(i) sum(i >= mondays))
datesmonths <- data.frame(months = months, timeSinceFirstDay = as.numeric(months - firstDate))

```

## Population

Source UK, ONS: <https://www.ons.gov.uk/file?uri=%2fpeoplepopulationandcommunity%2fpopulationandmigration%2fpopulationestimates%2fdatasets%2fpopulationestimatesforukenglandandwalesscotlandandnorthernireland%2fmid2020/ukpopestimatesmid2020on2021geography.xls>

Source France, INSEE: <https://www.insee.fr/fr/statistiques/2381472>

Both grouped by age class

```{r}
# France
popFR <- read.csv("data/popFrance_insee.csv")
# Turn into dictionary
popFr <- popFR$pop
# Rewrite age classes
agcl <- matrix(unlist(strsplit(popFR$ageClass, "_")), byrow = TRUE, ncol = 2)[, 2]
agcl[agcl == "120"] <- "80"

names(popFr) <- as.numeric(agcl)

# Add pop information to France data
datFR$pop <- popFr[as.character(datFR$clage_vacsi)]
table(datFR$pop, useNA = "ifany")

# Relative number of doses
datFR$p1 <- datFR$d1 / datFR$pop
datFR$p2 <- datFR$d2 / datFR$pop
datFR$p3 <- datFR$d3 / datFR$pop


# England
popEN <- read.csv("data/ONS_popEstimates_mid2020.csv")
# Turn into dictionary 
popEn <- popEN$pop
names(popEn) <- popEN$ageClass

datEN$pop <- popEn[datEN$age]
table(datEN$pop, useNA = "ifany")

# Relative number of doses
datEN$p1 <- datEN$d1 / datEN$pop
datEN$p2 <- datEN$d2 / datEN$pop
datEN$p3 <- datEN$d3 / datEN$pop
```

## Aggregate

```{r}
cols <- c("d1", "d2", "d3", "p1", "p2", "p3")

datFRwk <- aggregate(datFR[, cols], by = list(wk = datFR$wk, clage_vacsi = datFR$clage_vacsi), FUN = sum)

datENwk <- aggregate(datEN[, cols], by = list(wk = datEN$wk, age = datEN$age), FUN = sum)

```


### Ages on aggregated data

```{r}
maxAge <- 95

# EN
# Extract min and max age
datENwk$ageMin <- substr(datENwk$age, 1, 2)
datENwk$ageMax <- substr(datENwk$age, 4, 5)
datENwk[datENwk$age == "90+", "ageMax"] <- as.character(maxAge)

# Turn into numeric
datENwk$ageMin <- as.numeric(datENwk$ageMin)
datENwk$ageMax <- as.numeric(datENwk$ageMax)


# FR
# Remove all ages
datFRwk <- datFRwk[datFRwk$clage_vacsi > 0, ]
sort(unique(datFRwk$clage_vacsi))

datFRwk$ageMax <- datFRwk$clage_vacsi
datFRwk[datFRwk$ageMax == 80, "ageMax"] <- maxAge

datFRwk$ageMin <- NA
datFRwk[datFRwk$clage_vacsi == 4, "ageMin"] <- 0
datFRwk[datFRwk$clage_vacsi == 9, "ageMin"] <- 5
datFRwk[datFRwk$clage_vacsi == 11, "ageMin"] <- 10
datFRwk[datFRwk$clage_vacsi == 17, "ageMin"] <- 12
datFRwk[datFRwk$clage_vacsi == 24, "ageMin"] <- 18
datFRwk[datFRwk$clage_vacsi == 29, "ageMin"] <- 25
datFRwk[datFRwk$clage_vacsi == 39, "ageMin"] <- 30
datFRwk[datFRwk$clage_vacsi == 49, "ageMin"] <- 40
datFRwk[datFRwk$clage_vacsi == 59, "ageMin"] <- 50
datFRwk[datFRwk$clage_vacsi == 64, "ageMin"] <- 60
datFRwk[datFRwk$clage_vacsi == 69, "ageMin"] <- 65
datFRwk[datFRwk$clage_vacsi == 74, "ageMin"] <- 70
datFRwk[datFRwk$clage_vacsi == 79, "ageMin"] <- 75
datFRwk[datFRwk$clage_vacsi == 80, "ageMin"] <- 80

table(datFRwk$ageMax, useNA = "ifany")
table(datFRwk$ageMin, useNA = "ifany")
```

### Dates on aggregated data

```{r}
tmp <- sapply(mondays, FUN = function(i) sum(i >= mondays))
dates <- data.frame(wk = tmp, mondays = mondays, timeSinceFirstDay = as.numeric(mondays - firstDate))

datFRwk <- merge(datFRwk, dates, by = "wk", all.x = TRUE)
datENwk <- merge(datENwk, dates, by = "wk", all.x = TRUE)
```

# Plot


```{r countriesDoses, fig.width=12, fig.height=5}

for(ctr in c("EN", "FR")){
    
  
  plotPNG <- FALSE # Whether to save as png (no need if Knit)
  fname <- paste0("pics/joyPlot_", ctr, ".png") # Fig file name
  if(plotPNG) png(fname, width = 1000, height = 400)
  
  # Get the data for the country we want
  dat <- get(paste0("dat", ctr, "wk"))
  
  # Find the maximum p value for scaling
  maxp <- max(c(dat$p1, dat$p2, dat$p3))
  # Function to scale the colors
  scalefunc <- function(p, exponent = 1) (p/maxp)^(exponent)
  
  
  # Initialize plot window
  par(las = 1)
  par(mar = c(4, 2.5, 4, 2.5))
  par(mgp = c(2, 0, 0))
#  par(bg = "black", fg = gray(0.8))
  par(bg = "white", fg = "black")
  
  # Plot parameters
  # Colors for doses
  color1 <- "#FFC107" # Alternative: "#004D40"
  color2 <- "#1E88E5"
  color3 <- "#D81B60"
  # Color for border of the rectangles
  colBorder <- par("bg")
  lwdd <- 0.2 # width of the border

  plot(NULL, 
       ylim = c(0, maxAge), 
       xlim = c(0, max(dat$timeSinceFirstDay)+14), 
       xlab = "", ylab = "", 
       xaxs = "i", yaxs = "i", frame.plot = FALSE, axes = FALSE)
  
  # For each week
  for (wkk in sort(unique(dat$wk))){
    
    # Subset of the data for this week
    m <- dat[dat$wk == wkk, ]
  
    # First doses
    rect(xleft = m$timeSinceFirstDay, 
         xright = m$timeSinceFirstDay + 7, 
         ybottom = m$ageMin, 
         ytop = m$ageMax + 1, 
         border = colBorder, 
         col = sapply(m$p1, function(i) adjustcolor(color1, alpha.f = scalefunc(i))), 
         lwd = lwdd)
    
    # Second doses
    rect(xleft = m$timeSinceFirstDay, 
         xright = m$timeSinceFirstDay + 7, 
         ybottom = m$ageMin, 
         ytop = m$ageMax + 1, 
         border = colBorder, 
         col = sapply(m$p2, function(i) adjustcolor(color2, alpha.f = scalefunc(i))), 
         lwd = lwdd)
    
    # Third doses
    rect(xleft = m$timeSinceFirstDay, 
         xright = m$timeSinceFirstDay + 7, 
         ybottom = m$ageMin, 
         ytop = m$ageMax + 1, 
         border = colBorder, 
         col = sapply(m$p3, function(i) adjustcolor(color3, alpha.f = scalefunc(i))), 
         lwd = lwdd)
  
  }
  
  # Annotations
  # Axes, vertical
  cexx <- 0.8
  axis(2, lwd = 0, at = seq(0, maxAge, by = 10), col.axis = par("fg"), cex.axis = cexx)
  axis(4, lwd = 0, at = seq(0, maxAge, by = 10), col.axis = par("fg"), cex.axis = cexx)
  mtext("age", side = 2, line = 1.5, las = 3, cex = cexx)

  # Horizontal axis
  axis(1, at = datesmonths$timeSinceFirstDay, 
       labels = format(datesmonths$months, "%b\n%Y"), cex.axis = cexx, padj = 0.75, lwd = 0, lwd.ticks = 1, tck = -0.01, col.axis = par("fg"))
  
  # Legend  
  legend(0, maxAge, inset = c(0, -0), col = c(color1, color2, color3), pch = 15, legend = c("Première Dose", "Dose 2 / complète", "Dose de rappel"), horiz = TRUE, yjust = 0, xpd = TRUE, bty = "n")
  
  # Title
  if(ctr == "EN"){
    pays <- "Angleterre"
  }else{
    pays <- "France"
  }
  mtext(pays, side = 3, line = 2, font = 2, cex = 1.3)
  
  # Credits
  if(ctr == "EN"){
    src <- "https://api.coronavirus.data.gov.uk/v2/data?areaType=nation&areaCode=E92000001&metric=vaccinationsAgeDemographics&format=csv"
  }else{
    src <- "https://www.data.gouv.fr/en/datasets/donnees-relatives-aux-personnes-vaccinees-contre-la-covid-19-1/"
  }
  mtext(side = 1, text = paste0("@flodebarre, inspiré de @VictimOfMaths, ", Sys.Date(), "\nDonnées : ", src, "\nCode : ", "https://github.com/flodebarre/covid_vaccination/blob/main/campagne_vaccinale.Rmd"), col = par("fg"), line = 3, family = "mono", adj = 0, cex = 0.5)
  
  if(plotPNG){
    dev.off()
    system(paste("open", fname))
  }
  
}

# Append the output
system("convert campagne_vaccinale_files/figure-html/countriesDoses-* -append campagne_vaccinale_files/figure-html/countriesDoses.png")
  
```



