---
title: "comparison_departements"
author: "FD"
output: 
  html_document: 
      code_folding: hide
      toc: TRUE
      toc_float: TRUE
      self_contained: no
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

```{r, include = FALSE}
dlData <- FALSE
# Whether to download the data again
```


# Load data

## Vaccination par département

Source <https://datavaccin-covid.ameli.fr/explore/dataset/donnees-vaccination-par-tranche-dage-type-de-vaccin-et-departement/information/?sort=-date_reference>

```{r loadData}
URL.AmeliDep <- "https://datavaccin-covid.ameli.fr/explore/dataset/donnees-vaccination-par-tranche-dage-type-de-vaccin-et-departement/download/?format=csv&timezone=Europe/Berlin&lang=fr&use_labels_for_header=true&csv_separator=%3B"
## does not work

dataFile <- paste0("data/donnees-vaccination-par-tranche-dage-type-de-vaccin-et-departement_", Sys.Date(),".csv") # name file with today's date
if(dlData){
  system(paste0("wget -O ", dataFile, " ", URL.AmeliDep))
}

dat.residence <- read.csv(dataFile, sep = ";", stringsAsFactors = FALSE, dec = ",")
head(dat.residence)
```


## Vaccination par département d'injection

Source <https://www.data.gouv.fr/en/datasets/donnees-relatives-aux-personnes-vaccinees-contre-la-covid-19-1/>  
`vacsi-dep`

```{r}
URL <- "https://www.data.gouv.fr/en/datasets/r/4f39ec91-80d7-4602-befb-4b522804c0af"
URL.SPFadep <- "https://www.data.gouv.fr/en/datasets/r/83cbbdb9-23cb-455e-8231-69fc25d58111"

dataFile <- paste0("data/vacciDeps_injection_", Sys.Date(),".csv") # name file with today's date
dataFile2 <- paste0("data/vacciDeps_injection-a_", Sys.Date(),".csv") # name file with today's date
if(dlData){
  download.file(URL, dataFile) # download file from repo
  download.file(URL.SPFadep, dataFile2) # download file from repo
}
dat.injection <- read.csv(dataFile, sep = ";", stringsAsFactors = FALSE, dec = ",")
dat.injection.a <- read.csv(dataFile2, sep = ";", stringsAsFactors = FALSE, dec = ",")
head(dat.injection)
head(dat.injection.a)
```

## Clean data

```{r, results = 'hide'}
# 
# ## Actually no real need to convert to numeric??
# 
# # Departement as numeric
# unique(dat.residence$departement_residence)
# dat.residence$dep_res <- dat.residence$departement_residence
# dat.residence[which(is.element(dat.residence$departement_residence, c("2A", "2B"))), "departement_residence"] <- "20" # Corse 
# 
# # Remove "Tout department" because is redundant
# dat.residence[which(is.element(dat.residence$departement_residence, c("Tout département"))), "departement_residence"] <- NA # Redundant
# 
# 
# dat.residence$dep <- as.numeric(dat.residence$departement_residence)
# sort(unique(dat.residence$dep))
# 
# unique(dat.injection$dep)
# dat.injection[which(is.element(dat.injection$dep, c("2A", "2B"))), "dep"] <- "20" # Corse 
# dat.injection$dep <- as.numeric(dat.injection$dep)
# head(dat.injection)
# unique(dat.injection$dep)
```


```{r}
# Compare age classes
# For the moment, we just look at all ages, but note that we could look at the detail as well one day when we find the time
sort(unique(dat.residence$classe_age))

sort(unique(dat.injection.a$clage_vacsi))
```


```{r, results = 'hide'}
# Select all ages and all vaccines


dat.residence <- dat.residence[which(dat.residence$classe_age == "TOUT_AGE" & dat.residence$type_vaccin == "Tout vaccin" & !is.na(dat.residence$dep)), ]

head(dat.residence)
# 999 : pas connu -> 0 (same code as other dataset)
#dat.residence[which(dat.residence$dep == 999), "dep"] <- 0

# Merge the datasets
# Keep only common dates and departements
dat <- merge(dat.residence, dat.injection, by.x = c("date", "departement_residence"), by.y = c("jour", "dep"))
head(dat)
sort(unique(tmp$date))

```

Dates

```{r}
dat.residence$date <- as.Date(dat.residence$date)
dat.injection$date <- as.Date(dat.injection$jour)
```

# Plots

## Initial versions

```{r}
colInjection <- "#e9a3c9"
colResidence <- "#a1d76a"
```

Consistency check

```{r}
# Find latest common date between the two datasets
finalCommonDate <- min(c(max(dat.injection$date), max(dat.residence$date)))

# Check another date
# finalCommonDate <- "2021-05-30"

final.residence <- dat.residence[which(dat.residence$date == finalCommonDate), ]
final.injection <- dat.injection[which(dat.injection$date == finalCommonDate), ]

unique(final.residence$dep)
unique(final.injection$dep)

rbind(sum(final.injection$n_cum_complet)
, sum(final.residence$effectif_cumu_termine))

# Compare numbers of injections in the two datasets
sum(final.injection$n_cum_dose1) - sum(final.residence$effectif_cumu_1_inj)
```

There are more injections in the by injection dataset than in the by residency dataset. So overall our difference will be positive (while should be 0). 

Merge datasets

```{r}
# Merge datasets
names(final.residence)
final.both <- merge(final.residence, final.injection, by = "dep", all = FALSE)

# Rename columns to make sure information about whether residence of injection is kept
final.both$cum1D_residence <- final.both$effectif_cumu_1_inj
final.both$cumTermine_residence <- final.both$effectif_cumu_termine
final.both$cum1D_injection <- final.both$n_cum_dose1
final.both$cumTermine_injection <- final.both$n_cum_complet

# Replace NA by 0
for(i in c("cum1D_residence", "cumTermine_residence", "cum1D_injection", "cumTermine_injection")){
  final.both[is.na(final.both[, i]), i] <- 0
}

# Difference between nb injections and nb residence
final.both$diffInjRes.1D <- final.both$cum1D_injection - final.both$cum1D_residence

final.both$diffInjRes.termine <- final.both$cumTermine_injection - final.both$cumTermine_residence

# Relative difference
final.both$reldiffInjRes.1D <- (final.both$cum1D_injection - final.both$cum1D_residence)/final.both$cum1D_residence

final.both$reldiffInjRes.termine <- (final.both$cumTermine_injection - final.both$cumTermine_residence)/final.both$cumTermine_residence

# Consistency check
sum(final.both$diffInjRes.1D) # NB not the same as before because more geog info in res; to get the same number, add "all = TRUE" in the merge function
```

Plot differences per departement 
```{r}
relatif <- TRUE

if(relatif){
  txt <- "relative"
}else{
  txt <- ""
}

filename <- paste0("injections-residents", txt,".png")
png(filename = filename, width = 1600, height = 500, res = 100, pointsize = 12)

par(xpd = FALSE)
# Column to be plotted
if(relatif){
  thecol <- "reldiffInjRes.1D"
}else{
  thecol <- "diffInjRes.1D"
}

# Sort data by this columns
tmp <- final.both[order(final.both[, thecol], decreasing = TRUE), ]
# Number of different departments
n <- nrow(tmp)

par(las = 1)
par(mar = c(1, 5, 3, 5))

stopifnot(tmp[n, "dep"] == 0) # Check that last one corresponds to data for which dep info was not known (initially coded 0 or 990 in the different datasets)

tmpp <- tmp[-n, thecol] # Data for which dep info is known
ymax <- max(tmpp[which(tmpp<Inf)]) # max y value

x0 <- n+6 # Position of the point for which dep info is not known
y0 <- 2.5*min(tmpp) # y position of the point (broken y axis)

# Initialize plot
plot(tmp[-n, thecol], xlim = c(1, x0), ylim = c(y0, ymax), axes = FALSE, 
     xlab = "", ylab = "", 
     type = "n")
mtext(paste0("Différence ", txt), side = 2, las = 3, line = 3)
mtext(paste0("Différence ", txt, " entre nombre d'injections dans le département
et nombre de personnes vaccinées habitant le département"))
axis(2)
usr <- par("usr")

# Position of the last point (no geog info)
xpos4 <- x0 + 4
axis(4, at = c(0, round(y0, 2)), pos = xpos4) # Add right axis

# Add horizontal lines
clip(usr[1], n, usr[3], usr[4])
for(i in seq(-0.3, 0.6, by = 0.1)){
  abline(h = i, lwd = 0.5, col = gray(0.5))
}
clip(usr[1], usr[2], usr[3], usr[4])
abline(h = 0, lwd = 1.5, col = gray(0.5))


# Add symbols for broken axis
par(xpd = TRUE)
ygap <- y0*3/4
points(x = c(xpos4, n + (x0 - n)/2), y = c(ygap, 0), cex = 3, col = "white", pch = 15)
text(x = xpos4, y = ygap, labels = "//", cex = 1.5)

# Define colors for positive and negative values
colPos <- "#f1a340"
colNeg <- "#998ec3"
cols <- c(colNeg, colPos)
names(cols) <- c("Neg", "Pos")

# Add points
points(tmp[-n, thecol], col = cols[1 + (tmp[-n, thecol] > 0)], pch = 16)
points(x0, y0, col = cols[1 + (tmp[n, thecol] > 0)], pch = 16) # last one

# Add legend, dep info
text(x = c(1:(n-1), x0), y = c(tmp[-n, thecol], y0), labels = c(tmp[-n, "dep"], "pas de dep."), cex = 0.55, adj = c(0.5, -1))

# Identify position at which change from pos to neg values happens
xposneg <- which(tmp[-n, thecol] < 0)[1] - 1/2

# Add legend for signs
text(x = c(1 + xposneg/2, xposneg + (x0 - xposneg)/2), y = 0.65*c(usr[4], usr[3]), labels = c("Plus de personnes primo-injectées dans le département (données SPF)\nque de résidents du département primo-injectés (données Ameli)", 
"Moins de personnes primo-injectées dans le département (données SPF)\nque de résidents du département primo-injectés (données Ameli)"), col = c(colPos, colNeg))

dev.off()
system(paste("open", filename))
```

## New version

```{r}
# Compute differences in couverture among datasets
# 1 injection
dat$diff_1_inj <- -(as.numeric(dat$taux_cumu_1_inj) * 100 - as.numeric(dat$couv_dose1))

# full
dat$diff_complet <- -(as.numeric(dat$taux_cumu_termine) * 100 - as.numeric(dat$couv_complet))

# Maximum absolute value (to scale the figures)
rangemax <- range(c(dat$diff_1_inj, dat$diff_complet), na.rm = TRUE)
dmax <- max(abs(rangemax))
dmax
dmax.round <- ceiling(dmax/10)*10
```

```{r}
dept <- "92"
sub <- dat[dat$departement_residence == dept,]
head(sub)

plot(as.Date(sub$date), as.numeric(sub$taux_cumu_1_inj) * 100, ylim = c(0, 100))
points(as.Date(sub$date), sub$couv_dose1)


plot(as.Date(sub$date), as.numeric(sub$taux_cumu_1_inj) * 100, ylim = c(0, 100))
points(as.Date(sub$date), sub$couv_dose1)

plot(as.Date(sub$date), as.numeric(sub$effectif_cumu_1_inj))
points(as.Date(sub$date), sub$n_cum_dose1)

xx <- as.Date(sub$date)
yy <- as.numeric(sub$taux_cumu_1_inj) * 100 - as.numeric(sub$couv_dose1)
ipos <- (yy > 0)
izero <- (yy == 0)
ineg <- (yy < 0)

colPos <- "red"
colNeg <- "blue"
colZero <- "black"
lwdh <- 6
par(las = 1)
plot(xx, yy, ylim = c(-dmax, dmax), 
     xlab = "Date", 
     ylab = "Ameli - SPF, % points", 
     type = "n")
points(xx[ipos], yy[ipos], col = colPos, type = "h", lend = "butt", lwd = lwdh)
points(xx[ineg], yy[ineg], col = colNeg, type = "h", lend = "butt", lwd = lwdh)
points(xx[izero], yy[izero], col = colZero, type = "h", lend = "butt", lwd = lwdh)
abline(h = 0)

```

## Geographic

## Geographic

### L452R 

Source idea map: [Le Monde map](https://www.lemonde.fr/les-decodeurs/article/2020/05/05/coronavirus-age-mortalite-departements-pays-suivez-l-evolution-de-l-epidemie-en-cartes-et-graphiques_6038751_4355770.html) [| archived](https://archive.is/eJ68m)

```{r}
# Colors
colPos <- "#00918C" # Diff value is greater than the + threshold
colNeg <- "#EE9D00" # Diff value is smaller than the - threshold
colZero <- gray(0.75) # Diff value is between the - and + threshold
lwdh <- 6

par(las = 1)

# Function to scale values
scale.xy <- function(z, z1, z2, yrange){
  # z raw values
  # z1 new min value after scaling
  # z2 new max value after scaling
  # yrange range of values of z (if goes beyong z)
  z01 <- (z - yrange[1])/(yrange[2] - yrange[1])
  # z01 has to be between 0 and 1
  stopifnot(z01 >= 0 | z01 <= 1 | is.na(z01))
  z1 + (z2 - z1) * z01
}

# Function to draw single rectangle figure
drawRec <- function(depDat, x, y, dxy, whichDiff = "1_inj", yrange, thrCol){
  # depDat: dataset for this departement
  # x: x position of the bottom left corner
  # y: y position of the bottom left corner
  # dxy: rectangle size c(dx, dy)
  # whichDiff: "1_inj" or "complet", which metric to plot 
  # yrange: range of y values
  # thrCol: threhold absolute difference value, above which colorize

  # x values, time
  relTime <- as.numeric(as.Date(depDat$date) - min(as.Date(depDat$date)))/as.numeric(max(as.Date(depDat$date)) - min(as.Date(depDat$date)))
  # y values, couverture
  diffCouv <- depDat[, paste0("diff_", whichDiff)]
  
  # Scale x and y values  
  xx <- scale.xy(relTime, x, x + dxy[1], yrange = c(0, 1))
  yy <- scale.xy(diffCouv, y, y + dxy[2], yrange = yrange)
  
  # Identify indices depending on the chosen threshold
  ipos <- (diffCouv > thrCol)
  izero <- (diffCouv <= thrCol & diffCouv >= -thrCol)
  ineg <- (diffCouv < -thrCol)

  # Plot values separately to have different colors
  points(xx[ipos], yy[ipos], col = colPos, pch = 16, cex = 0.2, lwd = 0.7)
  points(xx[ineg], yy[ineg], col = colNeg, pch = 16, cex = 0.2, lwd = 0.7)
  points(xx[izero], yy[izero], col = colZero, pch = 16, cex = 0.2, lwd = 0.7)
  
  
}
```

```{r, fig.width = 7, fig.height = 9}
# Load positions of the departements
geog <- read.csv("../nouveauCriblage/data/position_deps.csv", header = FALSE)
names(geog) <- c("x", "y", "dep", "shortName")
rr <- 2 # Rounding factor

# Extract sizes to get rectangle size
d1 <- c(5.367367157581877, 1.008996513422371)
d2 <- c(6.203498009030808, 1.7266142092327605)
dd <- d2 - d1
ddr <- round(dd, rr)

thrCol <- 5 # Threshold absolute value of the difference for coloring
ylm <- rangemax # ylim for the rectangles # c(-dmax, dmax)

par(mar = rep(0.2, 4) + c(2, 0, 2, 0))
# Initialize plot with rect positions
plot(c(round(geog$x, rr), round(geog$x, rr) + ddr[1]), c(round(geog$y, rr), round(geog$y, rr) + ddr[2]), type = "n", asp = 1, axes = FALSE, xlab = "", ylab = "")

# Add titles of the plots
text(x = round(geog$x, rr) + ddr[1]/2, y = round(geog$y, rr) + ddr[2], labels = paste0(geog$shortName, "(", geog$dep, ")"), bg = "white", adj = c(0.5, -0.4), cex = 0.45)

# For each departement
for(i in seq_len(nrow(geog))){
  # Draw rectangle
  rect(xleft = round(geog[i, "x"], rr), ybottom = round(geog[i, "y"], rr), xright = round(geog[i, "x"], rr) + ddr[1], ytop = round(geog[i, "y"], rr) + ddr[2], lwd = 0.5, border = gray(0.6))
  
  # Select departement data
  depDat <- dat[which(dat$departement_residence == geog[i, "dep"]), ]
  
  # Horizontal line for 0 
  lines(c(round(geog[i, "x"], rr), round(geog[i, "x"], rr) + ddr[1]), rep(scale.xy(0, round(geog[i, "y"], rr), round(geog[i, "y"], rr) + ddr[2], yrange = ylm), 2), lty = 1, col = gray(0.6), lwd = 0.5)
  
  # Draw curve
  drawRec(depDat, x = round(geog[i, "x"], rr), y = round(geog[i, "y"], rr), dxy = ddr, whichDiff = "1_inj", yrange = ylm, thrCol = thrCol)

#    readline(prompt="Press [enter] to continue")

}


#drawRec(depDat, x = 5.73820184045573, y = 7.01837788404748, dxy = ddr, col = col452R)

minDay <- format(as.Date(min(dat$date)), "%d/%m")
maxDay <- format(as.Date(max(dat$date)), "%d/%m")

# Position of Paris
xP <- 1.34031283936975
yP <- 15.0106959818622
cexLeg <- 0.45
text(x = xP, y = yP, adj = c(1, 0), labels = ylm[1], cex = 0.4)
text(x = xP, y = yP + ddr[2], adj = c(1, 0.5), labels = ylm[2], cex = cexLeg)
text(x = xP, y = scale.xy(0, yP, yP + ddr[2], ylm), adj = c(1, 0.5), labels = "0", cex = cexLeg)

text(x = c(xP, xP + ddr[1]), y = c(yP, yP), adj = c(0.5, 1.5), labels = c(minDay, maxDay), cex = cexLeg)

title("Difference entre données par lieu de vaccination (SPF) et lieu de résidence (Ameli)")

  mtext("  @flodebarre
  Données SPF https://www.data.gouv.fr/en/datasets/donnees-relatives-aux-personnes-vaccinees-contre-la-covid-19-1/#
  Données Ameli https://datavaccin-covid.ameli.fr/explore/dataset/donnees-vaccination-par-tranche-dage-type-de-vaccin-et-departement/information/
  Code : https://github.com/flodebarre/
  Idée carte : Les Decodeurs, Le Monde https://tinyurl.com/carteFRDecLM", side = 1, line = 1, cex = 0.6, col = gray(0.5), adj = 0)
  
cexLegend <- 0.6
legend("topright", legend = c("Proportion L452R, < 50% le dernier jour", "≥ 50% le dernier jour"), cex = cexLegend, bty = "n", col = c(col452R, col452R.complement), lty = 1, pch = 16)
legend("topright", legend = c("", paste0("Données tracées si au moins ", 100*thrp, "% des cas criblés
et au moins ", thrntot, " criblages sur la semaine. 
Itv confiance sur nb crib. moyen 7 derniers jours.")), cex = cexLegend, bty = "n")
```


# Export
```{r}
save(final.both, file = paste0("data/outputComparisonDeps_finalCommonDate.Rdata"))
```

