---
title: "Livraisons - vaccination"
author: "FD"
output: 
  html_document: 
      code_folding: hide
      toc: TRUE
      toc_float: TRUE
      self_contained: no
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

```{r, include = FALSE}
dlData <- FALSE
# Whether to download the data again
```

# Initializations

# Load data and clean data

## Livraisons

Source <https://www.data.gouv.fr/en/datasets/donnees-relatives-aux-livraisons-de-vaccins-contre-la-covid-19/>

```{r, results = "hide"}
URL_livraisons <- "https://www.data.gouv.fr/en/datasets/r/c04da7da-be58-450e-bf3e-5993ce7796d9"

dataFile <- "data/livraisons.csv"
if(dlData){
  download.file(URL_livraisons, dataFile)
}
dat.livraisons <- read.csv(dataFile, stringsAsFactors = FALSE)
head(dat.livraisons)

dat.livraisons
```


```{r, results = "hide", fig.show='hide'}
unique(dat.livraisons$type_de_vaccin)

dat.livraisons.AZ <- dat.livraisons[dat.livraisons$type_de_vaccin == "AstraZeneca", ]
dat.livraisons.Js <- dat.livraisons[dat.livraisons$type_de_vaccin == "Janssen", ]
dat.livraisons.Md <- dat.livraisons[dat.livraisons$type_de_vaccin == "Moderna", ]
dat.livraisons.Pz <- dat.livraisons[dat.livraisons$type_de_vaccin == "Pfizer", ]

plot(as.Date(dat.livraisons.Pz$date_fin_semaine), dat.livraisons.Pz$nb_doses, type = "h", frame.plot = FALSE, 
     xlab = "Date fin de semaine", ylab = "Nb doses", 
     main = "Quality check")
points(as.Date(dat.livraisons.AZ$date_fin_semaine), dat.livraisons.AZ$nb_doses, col = 2)
points(as.Date(dat.livraisons.Js$date_fin_semaine), dat.livraisons.Js$nb_doses, col = 3)
points(as.Date(dat.livraisons.Md$date_fin_semaine), dat.livraisons.Md$nb_doses, col = 4)


names(dat.livraisons.AZ)[2:4] <- c("typeAZ", "nb_ucd_AZ", "nb_doses_AZ")
names(dat.livraisons.Js)[2:4] <- c("typeJs", "nb_ucd_Js", "nb_doses_Js")
names(dat.livraisons.Md)[2:4] <- c("typeMd", "nb_ucd_Md", "nb_doses_Md")
names(dat.livraisons.Pz)[2:4] <- c("typePz", "nb_ucd_Pz", "nb_doses_Pz")

alltimes <- data.frame("date_fin_semaine" = sort(unique(dat.livraisons$date_fin_semaine)))

newdat <- data.frame(matrix(0, ncol = 4, nrow = 1))
names(newdat) <- names(dat.livraisons)

for(tv in unique(dat.livraisons$type_de_vaccin)){
  print(tv)
  subdat <- dat.livraisons[dat.livraisons$type_de_vaccin == tv, ]
  subdat <- merge(alltimes, subdat, all = TRUE, by = "date_fin_semaine")
  subdat$type_de_vaccin <- tv
  # Find first 
  firstNonNA <- which(!is.na(subdat$nb_ucd))[1]
  if(firstNonNA > 1){
    subdat[1:(firstNonNA - 1), c("nb_ucd", "nb_doses")] <- 0
  }
  
  # Then loop on the lines -- maybe there's a more elegant way, but we have small enough datasets not to think this too hard
  for(iline in firstNonNA:nrow(subdat)){
    # If NA on the line, use value of the previous line
    if(is.na(subdat[iline, "nb_ucd"])){
      subdat[iline, c("nb_ucd", "nb_doses")] <- subdat[iline - 1, c("nb_ucd", "nb_doses")]
    }
  }
  print(subdat)
  newdat <- rbind(newdat, subdat)
}
# Remove first line put to create the dataframe
newdat <- newdat[-1, ]

dat.livraisons2 <- newdat

# Quality check
cols <- 1:4
names(cols) <- unique(dat.livraisons$type_de_vaccin)
points(as.Date(dat.livraisons2$date_fin_semaine), 
     dat.livraisons2$nb_doses, 
     col = cols[dat.livraisons2$type_de_vaccin], pch = 4)


doses.livraisons.total <- aggregate(dat.livraisons2$nb_doses, by = list("date_fin_semaine" = dat.livraisons2$date_fin_semaine), FUN = sum)

names(doses.livraisons.total)[2] <- "nb_doses"
```

## Personnes vaccinees

Source <https://www.data.gouv.fr/fr/datasets/donnees-relatives-aux-personnes-vaccinees-contre-la-covid-19-1/>

> Fichiers avec le nombre de personnes ayant reçu au moins une dose ou deux doses, par vaccin, arrêté à la dernière date disponible :
>
> vacsi-tot-v-fra-YYYY-MM-DD-HHhmm.csv (échelle nationale)
>
> Fichiers avec le nombre de personnes ayant reçu au moins une dose ou complètement vaccinées, arrêté à la dernière date disponible :
> 
> vacsi-tot-fra-YYYY-MM-DD-HHhmm.csv (échelle nationale)
> 
> 0 : Tous vaccins  
> 1 : COMIRNATY Pfizer/BioNTech  
> 2 : Moderna    
> 3 : AstraZeneka  
> 4 : Janssen  
> 

```{r, results = "hide"}
URL_v12 <- "https://www.data.gouv.fr/fr/datasets/r/b273cf3b-e9de-437c-af55-eda5979e92fc"

dataFile <- "data/vaccin_12doses.csv"
if(dlData){
  download.file(URL_v12, dataFile)
}
dat.vaccin12 <- read.csv(dataFile, stringsAsFactors = FALSE, sep = ";")
head(dat.vaccin12)

dat.vaccins <- dat.vaccin12[which(dat.vaccin12$vaccin != 0), ]

dic.typeVaccin <- c("Pfizer", "Moderna", "AstraZeneca", "Janssen")
names(dic.typeVaccin) <- as.character(1:4)
dat.vaccins$typeVaccin <- dic.typeVaccin[dat.vaccins$vaccin]

# Subselect tous vaccins
dat.vaccin12 <- dat.vaccin12[which(dat.vaccin12$vaccin == 0), ]

# Number of doses
dat.vaccin12$n_doses <- dat.vaccin12$n_dose1 + dat.vaccin12$n_dose2

dat.vaccin12$n_cum_doses <- cumsum(dat.vaccin12$n_doses)
```

```{r, eval = FALSE}
# This is actually not needed, but I am keeping it for legacy
URL_v1c <- "https://www.data.gouv.fr/fr/datasets/r/efe23314-67c4-45d3-89a2-3faef82fae90"

dataFile <- "data/vaccin_1cdoses.csv"
if(dlData){
  download.file(URL_v1c, dataFile)
}
dat.vaccin1c <- read.csv(dataFile, stringsAsFactors = FALSE, sep = ";")
head(dat.vaccin1c)
```

```{r, eval = FALSE}
par(las = 1)
plot(as.Date(dat.vaccin12$jour), dat.vaccin12$n_cum_dose1, type = "l", 
     xlab = "date", ylab = "number individuals")

lines(as.Date(dat.vaccin12$jour), dat.vaccin12$n_cum_dose2, type = "l", col = 2)

lines(as.Date(dat.vaccin1c$jour), dat.vaccin1c$n_cum_dose1, type = "l", col = 3, lty = 2)

lines(as.Date(dat.vaccin1c$jour), dat.vaccin1c$n_cum_complet, type = "l", col = 4, lty = 1)

legend("topleft", col = c(1, 2, 3, 4), lty = c(1, 1, 2, 1), legend = c("1 dose au moins", "2 doses", "1 dose au moins", "complet"), bty = "n")
```

We can deduce the number of people with complete vaccination after one dose (dose1_complet) and one dose and incomplete vaccination (but actually we do not need this for vaccine doses)

```{r, eval = FALSE}
# Just keeping this for legacy
# Quality checks
for(col in c("n_dose1", "n_cum_dose1")){
  print(all(dat.vaccin12[, col] == dat.vaccin1c[, col]))
}

dat.vaccin <- merge(dat.vaccin12, dat.vaccin1c[, c("jour", "n_complet", "n_cum_complet")], by = "jour")

# Other quality checks (cum sums)
all(dat.vaccin$n_cum_dose1 == cumsum(dat.vaccin$n_dose1))
all(dat.vaccin$n_cum_dose2 == cumsum(dat.vaccin$n_dose2))
all(dat.vaccin$n_cum_complet == cumsum(dat.vaccin$n_complet))

# 1 dose and complete
dat.vaccin$n_dose1complet <- dat.vaccin$n_complet - dat.vaccin$n_dose2
dat.vaccin$n_cum_dose1complet <- cumsum(dat.vaccin$n_dose1complet)

# 1 dose and incomplete
dat.vaccin$n_dose1incomplet <- dat.vaccin$n_dose1 - dat.vaccin$n_complet
dat.vaccin$n_cum_dose1incomplet <- cumsum(dat.vaccin$n_dose1incomplet)

# Quality check
all(dat.vaccin$n_cum_dose1 == dat.vaccin$n_cum_dose2 + dat.vaccin$n_cum_dose1complet + dat.vaccin$n_cum_dose1incomplet)

```


### Homogenize datasets

Function to add NA lines or fill in with info from previous date

```{r, results = "hide"}
# All dates
xx <- sort(unique(dat.vaccin12$jour))

fillTable <- function(mat, x = xx){
  # mat is the matrix with same columns as dat.vaccins, 
  # x is the full vector of dates; 
  #   the dates in the data must be included in x
  
  # Check that dates are sorted
  check1 <- all(mat[order(mat$jour), "jour"] == mat$jour)
  if(!check1) stop("make sure to order dates")
  
  # Check that no duplicated date
  if(any(duplicated(mat$jour))) stop("There are duplicated dates (maybe multiple vaccine types?)")
  
  # Check that all dates are included in x
  if(!all(is.element(unique(mat$jour), x))) stop("x does not contain all the dates of the dataset")
  
  # New dataset will all the dates that we want
  newmat <- merge(data.frame(jour = x), mat, by = "jour", all = TRUE) # Fill with NA the missing dates
  
  # Find first non NA value, fill previous lines with 0
  firstNonNA <- which(!is.na(newmat$n_dose1))[1]
  if(firstNonNA > 1){ # If there are lines before
    newmat[1:(firstNonNA - 1), c("n_dose1", "n_dose2", "n_cum_dose1", "n_cum_dose2")] <- 0
  }
  
  # Then loop on the lines -- maybe there's a more elegant way, but we have small enough datasets not to think this too hard
  for(iline in firstNonNA:nrow(newdat)){
    # If NA on the line, use value of the previous line
    if(is.na(newmat[iline, "n_dose1"])){
      newmat[iline, -c("jour")] <- newmat[iline - 1, -c("jour")]
    }
  }
  newmat
}

# Compute them for each vaccine type
dat.Pz <- fillTable(dat.vaccins[dat.vaccins$typeVaccin == "Pfizer", ])
dat.Md <- fillTable(dat.vaccins[dat.vaccins$typeVaccin == "Moderna", ])
dat.AZ <- fillTable(dat.vaccins[dat.vaccins$typeVaccin == "AstraZeneca", ])
dat.Js <- fillTable(dat.vaccins[dat.vaccins$typeVaccin == "Janssen", ])

# Re-bind the dataset
dat.vaccins <- rbind(dat.Pz, dat.Md, dat.AZ, dat.Js)
names(dat.vaccins) <- names(dat.Pz)

# Sum first dose and second dose to have the total number of injected doses
dat.vaccins$n_cum <- dat.vaccins$n_cum_dose1 + dat.vaccins$n_cum_dose2
```

# Combine datasets

## All vaccines

```{r}
col_injections <- "#2a9d8f"
col_doses <- "#e76f51"
lwdd <- 2

yy <- pretty(seq(0, max(doses.livraisons.total$nb_doses), 5))/10^6
# yy 
xx <- seq(as.Date("2020-12-01"), as.Date("2021-08-01"), by = "month")
# xx

par(mar = c(4.5, 4, 2, 4), las = 1)
plot(as.Date(doses.livraisons.total$date_fin_semaine), doses.livraisons.total$nb_doses/(10^6), type = "n", 
     xlab = "", 
     ylab = "", 
     col = col_doses, lwd = lwdd, 
     ylim = range(yy), 
     yaxs = "i", xaxs = "i", 
     frame.plot = FALSE)

for(i in seq(0, max(yy), by = 10)) abline(h = i, col = gray(0.9))

legend("topleft", col = c(col_doses, col_injections), lty = 1, legend = c("Total doses livrées", "Total doses injectées"), lwd = lwdd, title = "En millions:", box.lwd = 0.0, inset = c(0.02, 0), bg = gray(1, 0.5), bty = "o")



axis(1, at = xx, labels = format(xx, "%b"))
axis(2)

lines(as.Date(doses.livraisons.total$date_fin_semaine), doses.livraisons.total$nb_doses/(10^6), col = col_doses, lwd = lwdd)
lines(as.Date(dat.vaccin12$jour), dat.vaccin12$n_cum_doses/10^6, type = "l", col = col_injections, lwd = lwdd)

xmax <- max(c(as.Date(dat.vaccin12$jour), as.Date(doses.livraisons.total$date_fin_semaine)))

par(xpd = TRUE)
text(rep(as.Date(xmax), 2), c(max(dat.vaccin12$n_cum_doses/10^6), max(doses.livraisons.total$nb_doses/(10^6))), 
     labels = paste(round(c(max(dat.vaccin12$n_cum_doses/10^6), max(doses.livraisons.total$nb_doses/(10^6))), 1), "Mio"), 
     col = c(col_injections, col_doses), adj = c(0, 0.5))
par(xpd = FALSE)

mtext(side = 1, adj = 0, text = paste0(format(Sys.time(), "%Y-%m-%d"), " | Données : 
- Livraisons https://www.data.gouv.fr/en/datasets/donnees-relatives-aux-livraisons-de-vaccins-contre-la-covid-19/
- Injections https://www.data.gouv.fr/fr/datasets/donnees-relatives-aux-personnes-vaccinees-contre-la-covid-19-1/"), cex = 0.6, col = gray(0.5), line = 3)
```


## By type of vaccine

### Livraisons et injections 

```{r}
# Define colors of vaccine types
colType <- c("blue", "red", "orange", rgb(0, 0.6, 0))
names(colType) <- c("Pfizer", "Moderna", "AstraZeneca", "Janssen")

# Define pch of vaccine types
pchType <- 0:3
names(pchType) <- names(colType)

# Line width
lwdd <- 2

# Denominator for scale
# (1 for full value, 10^6 if by million)
denom <- 10^6

par(las = 1)
layout(matrix(1:4, ncol = 2))

  # For each vaccine type
  for(typ in names(colType)){
    # Subsets of the data with this type
  subliv <- dat.livraisons2[which(dat.livraisons2$type_de_vaccin == typ), ]
  subdoses <- dat.vaccins[dat.vaccins$typeVaccin == typ, ]
  
  # Set ymax value (same scale for smaller values)
  ymax <- 1.05 * max(subliv$nb_doses)
  if(ymax < 12*10^6) ymax <- 12*10^6
  
  par(mar = c(2, 3, 2, 2), 
      mgp = c(1.5, 0.25, 0), tck = -0.015, xpd = TRUE, 
      las = 1)
  
  plot(as.Date(subliv$date_fin_semaine), subliv$nb_doses/denom,
       col = colType[typ], pch = pchType[typ], 
       type = "o", 
       ylim = c(0, ymax/denom), 
       xlab = "", ylab = "en millions", main = typ, yaxs = "i")
  
  legend("topleft", col = colType[typ], pch = c(pchType[typ], NA), lty = 1, lwd = c(1, lwdd), legend = c("livraisons", "injections"), bty = "n", inset = c(0.01, 0))
  
  lines(as.Date(subdoses$jour), subdoses$n_cum/denom, col = colType[typ], lwd = lwdd, pch = pchType[typ])
  
  axis(4)
  par(xpd = FALSE)
}

```

### Ecart entre livraisons et injections 

```{r}
# Combine the two datasets to make sure we get the same dates
dat.combined <- merge(dat.livraisons2, dat.vaccins, by.x = c("date_fin_semaine", "type_de_vaccin"), by.y = c("jour", "typeVaccin"), all = TRUE)

# Compute remaining number of doses
dat.combined$remaining <- dat.combined$nb_doses - dat.combined$n_cum

par(mfrow = c(1, 1), xpd = FALSE, 
    mgp = c(1.25, 0.25, 0), tck = -0.01, 
    mar = c(4, 3, 3, 3), las = 1)

plot(as.Date(dat.combined$date_fin_semaine), dat.combined$remaining/denom, col = colType[dat.combined$type_de_vaccin], type = "n", yaxs = "i", 
     xlab = "", ylab = "en millions", 
     axes = FALSE, main = "Doses disponibles")

mtext("date fin semaine", side = 1, line = 3)

# x axis
xx <- sort(unique(dat.livraisons2$date_fin_semaine))
if(max(xx) > max(dat.vaccins$jour)){xx <- xx[1:(length(xx)-1)]}
axis(1, at = as.Date(xx), labels = format.Date(xx, "%Y-%m-%d"), las = 3, cex.axis = 0.6)

axis(2)
axis(4)

# Graduations
for(i in seq(0, 5, by = 0.5)){
  abline(h = i, col = gray(0.9))
}

cexx <- 0.75 # marker size
par(xpd = TRUE)
for(typ in names(colType)){
  # Subset of the data for this vaccine type
  subdat <- dat.combined[dat.combined$type_de_vaccin == typ, ]
  # Remove NAs
  # (injections by day, livraisons by week)
  # (necessary to plot the line between markers)
  subdat <- subdat[!is.na(subdat$remaining/denom), ]  
  
  points(as.Date(subdat$date_fin_semaine), subdat$remaining/denom, type = "o", lty = 1, col = colType[typ], pch = pchType[typ], cex = cexx)
}

legend("topleft", inset = c(0.01, 0), box.lwd = 0, legend = names(colType), lty = 1, col = colType, pch = pchType, pt.cex = cexx)

```

